Accurate binding of sodium and calcium to phospholipid bilayers by effective inclusion of electronic polarization

### 2017-11-08 Manuscript/manuscript.tex
Rewritten Abstract

\begin{abstract}
  % insert abstract here
  The binding affinities and stoichimetries of Na$^+$ and Ca$^{2+}$ ions to phospholipid bilayers have not been
  consistently determined with experimental and computational methods,
  despite their significance and biological relevance in cellular membranes. 
  The details concerning ion binding could be resolved with classical molecular dynamics (MD) simulations,
  however, the accuracy of the available lipid models has not been sufficient.
  In this work we show that the binding details of Na$^+$ and Ca$^{2+}$ ions to the
  1-Palmitoyl-2-oleoylphosphatidylcholine (POPC) bilayer
  can be accurately described with a MD simulation model implicitly including electronic polarization.
  This is demonstrated by applying the electronic continuum correction (ECC) to a
  state of the art lipid model for MD simulations of the POPC
  lipid bilayer. The newly introduced ECC-lipids model reproduces the experimentally measured structural parameters for the ion-free membrane,
  the response of lipid headgroup to a bound positive charge, and the binding affinities of
  Na$^+$ and Ca$^{2+}$ ions. The observed negligible binding of Na$^+$ ions to POPC bilayer
  and stronger interactions of Ca$^{2+}$ primarily with phoshpate oxygens give
  support to the early intepretations of the experimental spectroscopic data.
  At the same time, Ca$^{2+}$ ions form complexes with 1-3 lipid molecules with almost
  equal probabilities, suggesting a more complicated binding stoichimetry than 
  the ternary complex binding model used to interpret the NMR data.
  The results of this work pave the way to accurate MD simulations of complex biochemical systems
  with realistic electrostatic interactions in the vicinity of cellular membranes.
\end{abstract}

@@

Since artifacts were reported in simulations with \AA{}qvist ions is water \cite{auffinger07},
we also simulated the Lipid14 model with more realistic ion models by Dang et al.~\cite{smith94,chang1999,dang2006}, and
ECC-ions~\cite{jungwirth17-new-paper-to-be-published, Pluharova2014}. 
However, we still observed overestimated binding for both \ce{Ca^{2+}} and Na$^+$ with these ions models 
(Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}). 
The results are in line with the previous work \cite{catte16},
suggesting that additional improvements in the lipid parameters are
required to correctly describe the binding of cations to phospholipid bilayers.

@@

\begin{table}[tb!]
  \caption{Relative surface excess of calcium with respect to water, 
  and the ratios of the amount of calcium cations 
  bound to phosphate or carbonyl moieties 
  to the amount of such cations bound to both moieties in a POPC bilayer 
  from various simulation models. 
  $C_w$ is the concentration of CaCl$_2$ in water before adsorbing to the phospholipid bilayer. 
  \label{tab:binding}}
  \begin{tabular}{l|c | c c | c c}
    model                 & $\Gamma_{Ca}^w \,/\, \mathrm{nm}^{-2}$ & $C_{w}$ & $C_b\,/\,\mathrm{mM}$   & $r^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $r^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
    \hline
    ECC-lipids            &  $0.06 \pm 0.01 $ &  350  &  $280\pm 10 $  &  99\%  &    25\%    \\
    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $ &  350  &  $210\pm 10 $  & 100\%  &    37\%     \\
    Lipid14/Dang          &  $0.23 \pm 0.03 $ &  350  &  $160\pm 10 $  & 100\%  &    14\%    \\
    Lipid14/ECC-ions      &  $0.35 \pm 0.11 $ &  350  &  $120\pm 100$  & 100\%  &    23\%    \\
%Average no. of bound cations:
%L14+Aqvist: all:22.7, phos:22.7, carb:8.5
%L14+Dang:   all:29.2, phos:29.2, carb:4.0
%L14+ECC-ion:all:37.7, phos:37.7, carb:8.5
  \end{tabular}
\end{table}

>>>>

\begin{abstract}
  % insert abstract here
Binding affinities and stoichiometries of Na$^+$ and Ca$^{2+}$ ions to phospholipid
bilayers are of paramount significance in the properties and functionality of cellular
membranes. For example, possible accumulation of ions at the membrane interface determines
the interaction affinity with charged species. Current estimates of binding affinities and
stoichiometries of cations are controversial due to limitations in the experimental and
theoretical methods available. Experimentally one can assess these parameters by titrating
several membrane properties as a function of the ions concentration. However, the
interpretation of the obtained results requires theoretical models. Considering that the
real behavior is unknown these models are at best idealizations. Classical molecular
dynamics (MD) simulations provide details of the ions binding process with atomistic
resolution. Therefore, MD a priory offers all necessary information to interpret
experimental data without the need to resort to simplified models. However, the accuracy
of the available lipid models when interacting with ions is currently limited. In this
work, we improve the binding details of Na$^+$ and Ca$^{2+}$ ions to the
1-Palmitoyl-2-oleoyl-phosphatidylcholine POPC) bilayer by implicitly including electronic
polarization. We do this by applying a mean field correction to the partial charges known
as the electronic continuum correction (ECC) to a selected state of the art POPC lipid
model for MD simulations. Our improved ECC-POPC model reproduces not only the
experimentally measured structural parameters for the ion-free membrane, but also the
response of lipid headgroup to a bound positive charge, and the binding affinities of 
Na$^+$ and Ca$^{2+}$ ions. We also observe negligible binding of Na$^+$ ions to POPC
bilayer and stronger interactions of Ca$^{2+}$ primarily with phosphate oxygens in our new
model is in agreement with available experimental spectroscopic data. This new model
results in Ca$^{2+}$ ions forming complexes with one to three POPC molecules with almost
equal probabilities which suggests that previous models used to interpret NMR data are too
simple. The results of this work pave the way to quantitative MD simulations of complex
biochemical systems with realistic electrostatic interactions in the vicinity of cellular
membranes.
\end{abstract}

@@

Since artifacts were reported in simulations with \AA{}qvist ions is water \cite{auffinger07},
we also simulated the Lipid14 model with more realistic ion models by Dang et al.~\cite{smith94,chang1999,dang2006}, and
ECC-ions~\cite{jungwirth17-new-paper-to-be-published, Pluharova2014}. 
However, we still observed overestimated binding for both \ce{Ca^{2+}} and Na$^+$ with these ions models 
(Figs.~\ref{fig:delta_ordPar_CaCl} and~\ref{fig:delta_ordPar_NaCl}). 
\todo{Add OP-response of Lipid14+ECC-ions plot in SI}.
The results are in line with the previous work \cite{catte16},
suggesting that additional improvements in the lipid parameters are
required to correctly describe the binding of cations to phospholipid bilayers.

@@

\begin{table}[tb!]
  \caption{Relative surface excess of calcium with respect to water, 
  and the ratios of the amount of calcium cations 
  bound to phosphate or carbonyl moieties 
  to the amount of such cations bound to both moieties in a POPC bilayer 
  from various simulation models. 
  The CaCl$_2$ concentration of the ions in water before adsorption to the phospholipid bilayer
  is 350mM in the analyzed systems. \label{tab:binding}}
  \todo{Add columns with concentrations of calcium -- bulk/molar fraction.}
  \todo{Changing symbol $P$ to $R$ or $r$ might reflect better that it is rather a ratio, and not that much a probability.}
  \begin{tabular}{l|c c c}
    model                 & $\Gamma_{Ca}^w$ (nm$^{-2}$)   & $P^\mathrm{Ca^{2+}} _\mathrm{PO_4} $ & $P^\mathrm{Ca^{2+}} _\mathrm{O_{carb.}} $ \\
    \hline
    ECC-lipids            &  $0.06 \pm 0.01 $            &  99\%   &    25\%    \\
    Lipid14/\AA{}qvist    &  $0.13 \pm 0.01 $            &  -       &     -       \\
    Lipid14/Dang          &  $0.23 \pm 0.03 $            &  -        &    -       \\
    Lipid14/ECC-ions      &  $0.35 \pm 0.11 $            &  -        &    -       \\
  \end{tabular}
\end{table}
